Finding Concurrency Bugs in Production

### 2020-05-10 src/paper.tex
Small fixes for the first draft deadline

\subsection{The Go Programming Language}
All code examples are written in Go which is a statically typed programming language developed by Google, which compiles down to single native binaries.
The reason we are using Go and not C/C++ or Java is the recent popularity of it for writing highly concurrent programs.
Popular open-source projects like \emph{Kubernetes}~\footnote{\url{https://kubernetes.io/}}, \emph{Docker}~\footnote{\url{https://www.docker.com/}} or \emph{Prometheus}~\footnote{\url{https://prometheus.io/}} are all written in Go.
In the official documentation of Go it claims: ``Go is expressive, concise, clean, and efficient. Its concurrency mechanisms make it easy to write programs that get the most out of multicore and networked machines, while its novel type system enables flexible and modular program construction.''~\cite{goDocs}
So, Go tries to solve some of the problems that arise when writing concurrent software application by their design of language.
% TODO: How does Go want to archive this?
Go abstracts the creation of threads by using lightweight \emph{goroutines}, which can be created by using the \lstinline{go} followed by a named or anonymous function.
The Go runtime maps these goroutines to normal Kernel threads during execution.
Go encourages programmers to use concurrency by \emph{message passing}, which is assumed to be safer and more convenient to use than \emph{shared memory}.
However, it is also possible to make use of \emph{shared memory}, which gives the developer a lot of freedom to orchestrate the parallel execution of threads.
A study by Tu, Liu, Song and Zhan analyzed popular Go projects and their concurrency bugs.
They came to the conclusion that \emph{message passing} produced even more errors than \emph{shared memory}.
Projects written in Go also tend to have more concurrency than projects written in traditional programming languages such as C.~\cite{tu2019go}
Given all those conclusions it is even more important for Go developers to catch concurrency bugs effectively.
Together with the increasing popularity of the language, the demand for production ready tools to catch concurrency bugs is very high, which is why this paper evaluates existing techniques in the context of the Go programming language.

@@

A second example of a deadlock that might not be obvious is \Cref{lst:deadlockCh} which uses two unbuffered channels to transfer information between threads.
The problem here is that without an active listener on an unbuffered channel, any send action will be blocked.
To fix this, one could replace the unbuffered channel with a buffered one so that the execution flow of the program can continue without blocking.
This shows how important it is to know the concrete implementation of a concurrency abstraction.
Even though intended to ease the synchronization between threads and make inter-thread communication safer, by not knowing the implementation of an abstraction the developer can unknowingly create hard to find concurrency bugs.

@@

Another method of concurrency-aware testing has been proposed by Choi and Zeller.\cite{acm2002}
They implemented a tool called ``DEJAVU'' for the \emph{Jalapeño} Java Virtual Machine but the concept is also applicable to tools that work with Go or any other programming language.
They make use of a technique called \emph{Delta debugging}, which describes the process of narrowing down the spot where a failure is introduced by going back and forth between working and non-working conditions, in this case working and non-working thread schedules.~\cite{zeller2002delta}
\Cref{fig:testing} shows a modified model of the \emph{DEJAVU} approach, suited for Go programs by using rr instead of of \emph{DEJAVU} for the recording and replay of the thread schedule.
As a prerequisite, test cases need to be created that define the expected behavior of the application.
These can either succeed or fail which can be shown be the exit code for example.
In the first step, the Go application with these test cases gets executed and the thread schedule gets recorded by \emph{rr}.
The either failing or succeeding recorded thread schedule is then passed to the next instance where the delta debugging unit requests new executions of the application with an alternated thread schedule until it can pinpoint the failure-inducing thread switch.
The thread schedule is generated by the \emph{Schedule generator} and is replayed by the replay module in \emph{rr}.
The schedule as well as the test result is passed back to the delta debugging unit.
``By systematically narrowing down the difference between a thread schedule that makes the program pass and another schedule that makes the program fail, the Delta Debugging approach can pinpoint the error location automatically -- namely the location(s) where a thread switch causes the program to fail.''~\cite{acm2002}

>>>>

\subsection{The Go Programming Language}
All code examples are written in Go which is a statically typed programming language developed by Google, which compiles down to single native binaries.
The reason we are using Go and not C/C++ or Java is the recent popularity of it for writing highly concurrent programs.
Popular open-source projects like \emph{Kubernetes}~\footnote{\url{https://kubernetes.io/}}, \emph{Docker}~\footnote{\url{https://www.docker.com/}} or \emph{Prometheus}~\footnote{\url{https://prometheus.io/}} are all written in Go.
In the official documentation of Go it claims: ``Go is expressive, concise, clean, and efficient. Its concurrency mechanisms make it easy to write programs that get the most out of multicore and networked machines, while its novel type system enables flexible and modular program construction.''~\cite{goDocs}
So, Go tries to solve some of the problems that arise when writing concurrent software application by their design of language.
% TODO: How does Go want to archive this?
Go abstracts the creation of threads by using lightweight \emph{goroutines}, which can be created by using the \lstinline{go} keyword followed by a named or anonymous function.
The Go runtime maps these goroutines to normal Kernel threads during execution.
Go encourages programmers to use concurrency by \emph{message passing}, which is assumed to be safer and more convenient to use than \emph{shared memory}.
However, it is also possible to make use of \emph{shared memory}, which gives the developer a lot of freedom to orchestrate the parallel execution of threads.
A study conducted by Tu, Liu, Song and Zhan analyzed popular Go projects and their concurrency bugs.
They came to the conclusion that \emph{message passing} produced even more errors than \emph{shared memory}.
Projects written in Go also tend to have more concurrency than projects written in traditional programming languages such as C.~\cite{tu2019go}
Given all those conclusions it is even more important for Go developers to catch concurrency bugs effectively.
Together with the increasing popularity of the language, the demand for production ready tools to catch concurrency bugs is very high, which is why this paper evaluates existing techniques in the context of the Go programming language.

@@

A second example of a deadlock that might not be obvious is \Cref{lst:deadlockCh} which uses two unbuffered channels to transfer information between threads.
The problem here is that without an active listener on an unbuffered channel, any send action will be blocked.
To fix this, one could replace the unbuffered channel with a buffered one, so that the execution flow of the program can continue without blocking.
This shows how important it is to know the concrete implementation of a concurrency abstraction.
Even though intended to ease the synchronization between threads and make inter-thread communication safer, by not knowing the implementation of an abstraction the developer can unknowingly create hard to find concurrency bugs.

@@

Another method of concurrency-aware testing has been proposed by Choi and Zeller.\cite{acm2002}
They implemented a tool called ``DEJAVU'' for the \emph{Jalapeño} Java Virtual Machine but the concept is also applicable to tools that work with Go or any other programming language.
They make use of a technique called \emph{Delta debugging}, which describes the process of narrowing down the spot where a failure is introduced by going back and forth between working and non-working conditions, in this case working and non-working thread schedules.~\cite{zeller2002delta}
\Cref{fig:testing} shows a modified model of the \emph{DEJAVU} approach, suited for Go programs by using rr instead of of \emph{DEJAVU} for the recording and replay of the thread schedule.
As a prerequisite, test cases need to be created that define the expected behavior of the application.
These can either succeed or fail which can be shown by the exit code for example.
In the first step, the Go application with these test cases gets executed and the thread schedule gets recorded by \emph{rr}.
The either failing or succeeding recorded thread schedule is then passed to the next instance where the delta debugging unit requests new executions of the application with an alternated thread schedule until it can pinpoint the failure-inducing thread switch.
The thread schedule is generated by the \emph{Schedule generator} and is replayed by the replay module in \emph{rr}.
The schedule as well as the test result is passed back to the delta debugging unit.
``By systematically narrowing down the difference between a thread schedule that makes the program pass and another schedule that makes the program fail, the Delta Debugging approach can pinpoint the error location automatically -- namely the location(s) where a thread switch causes the program to fail.''~\cite{acm2002}
